---
title: "Regression in Biomedical Research"
author: "Jianing"
date: "2023-02-20"
output:
  html_document:
    df_print: paged
---
R has two native data formats—Rdata (sometimes shortened to Rda) and Rds. These formats are used when R objects are saved for later use. Rdata is used to save multiple R objects, while Rds is used to save a single R object.
```{r}
dt <- readRDS("/Users/jianingjin/Desktop/IEMS_304/lab4/Trp63.tf.rds")
head(dt)
library(glmnet)
library(Matrix)
dim(dt)
```

```{r}
#Part 1 LASSO Regression
#(1)
x <- model.matrix(Trp63~., data = dt)[, -1]
y <- dt$Trp63
#?
grid <- 10^seq(2, -2, length = 100)
set.seed(1)
train <- sample(1:nrow(x), nrow(x)/2) #x training
test <- (-train) #x testing
y.test <- y[test] #y testing
```

```{r}
set.seed(1)
cv.out <- cv.glmnet(x[train,], y[train], alpha = 1)
plot(cv.out)
```
```{r}
bestlamda <- cv.out$lambda.min
bestlamda
```
The which() function in R returns the position or the index of the value which satisfies the given condition. The which() function in R gives you the position of the value in a logical vector. The position can be of anything like rows, columns and even vector as well.
```{r}
#(2)
lasso.mod <- glmnet(x[train,], y[train], alpha = 1, lambda = bestlamda)
lasso.mod$beta[which(lasso.mod$beta != 0),] #beta: coefficients
```
```{r}
#(3)
lasso.pred <- predict(lasso.mod, s = bestlamda, x[test, ])
mean((lasso.pred - y.test)^2)
```

```{r}
#Part 2: Decision Trees
#Q1
library(tree)
dt2 = data.frame(dt) #alter dataset to dataframe
model.tree = tree(Trp63 ~ ., data = dt2, subset = train)
summary(model.tree)
```
```{r}
#Q2
plot(model.tree)
text(model.tree, pretty = 0)
```
The par() function is used to set or query graphical parameters. We can divide the frame into the desired grid, add a margin to the plot or change the background color of the frame by using the par() function. We can use the par() function in R to create multiple plots at once.
```{r}
#Q3
set.seed(1)
cv_out = cv.tree(model.tree)
par(mfrow = c(1, 2)) #function about lay out, makes plot in a same page
#plot(cv_out$dev, cv.out$size, type = "b")
#plot(cv_out$dev, cv.out$k, type = "b")
plot(cv_out$size, cv.out$dev, type = "b")
plot(cv_out$k, cv.out$dev, type = "b")
```
```{r}
which(cv_out$dev == min(cv_out$dev))
min(cv_out$dev)
```
```{r}
fit_prune <- prune.tree(model.tree, which(cv_out$dev == min(cv_out$dev)))
summary(fit_prune)
plot(fit_prune)
text(fit_prune, pretty = 0)
```

```{r}
pred_prune = predict(fit_prune, dt[-train, ])
test = dt[-train, "Trp63"]
plot(pred_prune, test)
abline(0, 1)
mean((pred_prune - test)^2)
```

```{r}
#Q4
library(randomForest)
set.seed(1)
fit_bag = randomForest(Trp63~., data = dt, subset = train, mtry = 225, importance = T)
```
```{r}
pred_bag = predict(fit_bag, dt[-train,])
plot(pred_bag, test)
abline(0, 1)
mean((pred_bag - test)^2)
```

```{r}
#Q5
set.seed(1)
fit_rf = randomForest(
  Trp63~., data = dt, subset = train, mtry = 76, importance = T
)
```

```{r}
pred_rf = predict(fit_rf, dt[-train, ])
mean((pred_rf - test)^2)
```

```{r}
#Q6
head(importance(fit_rf))
```
Conclusion
Q1: In reality, we would not focus only on one gene (e.g., Trp63) in our research. If we want to
extend one method to build a network including all other genes of our interests, from the results
above, which method would you choose to further apply? Why?
A: RandomForest. Beacuse this method has the smallest MSE.


Q2: After trying out all these regression methods, do you observe any consistency in the
regression results?xf What would you propose to be the most significant regulator genes for our gene
of interests Trp63?
A: Tfap2b
In Lasso regression, Tfap2b has the largest coefficient.
In Tree method, Tfap2b is in the first decision node every time.
In RandomForest, Tfap2b has the largest importance.